---
title: "Import data from 7-day experiment"
author: "Roy W. Martin"
date: "`r Sys.Date()`"
output:
  github_document:
    number_sections: TRUE
    df_print: "tibble"
    math_method: 
      engine: webtex
    #  url: https://latex.codecogs.com/svg.image?
    html_preview: TRUE
    keep_html: TRUE
editor_options:
  chunk_output_type: inline
  markdown: 
    wrap: 72
---

Below is code used to import data from an Excel file containing ICP-MS observations from a USEPA/ORD/CESER (Center for Environmental Solutions and Emergency Response - Cincinnati) laboratory experiment conducted over 7 days to characterize noise-limited or Ideal Background-Instrument Detection Limit (IB-IDL) for its local instrumentation.The observation-level data were measured as counts per second (CPS) recorded at specified m/z for individual "burns" from samples of 3 different $HNO_3$ 0.8% solutions. These original data were recorded to an MS Excel file, which was imported, modified, and organized into a data frame object within R for use in modeling in another script called "Import_ICPMS.Rmd" located at https://github.com/r-w-martin/ICPMS_7d_Exp. 
```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = TRUE)

library(readxl)
library(ggpubr)
library(moments)
library(ggplot2)
library(ggExtra)
library(gridExtra)
library(kableExtra)
library(ggrepel)
library(dplyr)
library(tidyverse)
library(tidyr)
library(future)
library(foreach)
library(itertools)
library(bayesplot)
library(tidybayes)
library(brms)

options(mc.cores = parallel::detectCores(logical = FALSE))
options( max.print = 1000 )
```


# Import data
Importing data from the main Excel spreadsheet provided by J. Creed USEPA/ORD/CESER.
```{r import, eval=FALSE, include=TRUE}
# set path and import all sheets and relevant columns from excel
path <- "./Seven Day individual replicates for Roy.xlsx"

df_import <- path %>%
  excel_sheets() %>%
  set_names() %>%
  map_df(~ read_excel(path = path, 
                      sheet = .x,
                      range = "A8:C220"),
         .id = "sheet")

# Remove NA column
# Add a column for burn number
# Rename columns with lower case headings
# Convert cps to integer
# Save original cps, but also multiply by 0.500 sec
# Convert sheet names to sample names

df_cleaned <- df_import %>%
  select(sheet, Mass, CPS) %>%
  mutate(burn = as.integer(substrRight(sheet, 1))) %>%
  rename(mass = Mass,
         cps = CPS) %>%
  mutate(cps_orig = cps,
         cps = as.integer(round(cps_orig)*0.5)) %>%
  mutate(sample = str_sub(sheet, 1, 5)) %>%
  mutate(day = case_match(sample,
                          "4BW32" ~ 1,
                          "4BW33" ~ 2,
                          "4BW34" ~ 3,
                          "4BW35" ~ 4,
                          "4BW36" ~ 5,
                          "4BW37" ~ 6,
                          "4BW38" ~ 7)) %>%
  mutate(sample_order = str_sub(sheet, 7, 8)) %>%
  mutate(description = case_match(sample_order,
                                  "01" ~ "0.8% HNO3 A",
                                  "02" ~ "ISO Solution 1 with Th",
                                  "03" ~ "ISO Solution 2",
                                  "04" ~ "Daily Solution New Na with 100ppb Er 5% HCL",
                                  "05" ~ "Multi Element",
                                  "06" ~ "0.8% HNO3 B",
                                  "07" ~ "0.4ppb Nd & Sm, 0.2ppb Gd, 50ppb Ce",
                                  "08" ~ "0.75ppb Nd & Sm, 0.375ppb Gd, 50ppb Ce",
                                  "09" ~ "1.25ppb Nd & Sm, 0.625ppb Gd, 50ppb Ce",
                                  "10" ~ "2.25ppb Nd & Sm, 1.125ppb Gd, 50ppb Ce",
                                  "11" ~ "3.75ppb Nd & Sm, 1.88ppb Gd, 50ppb Ce",
                                  "12" ~ "7.5ppb Nd & Sm, 3.75ppb Gd, 50ppb Ce",
                                  "13" ~ "15ppb Nd & Sm, 7.5ppb Gd, 50ppb Ce",
                                  "14" ~ "30ppb Nd & Sm, 15ppb Gd, 50ppb Ce",
                                  "15" ~ "0.8% HNO3 C",
                                  "16" ~ "0.1ppb Multi Element",
                                  "17" ~ "0.2ppb Multi Element",
                                  "18" ~ "Multi Element"))

save(df_cleaned, file = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/ICPMS_methods/ICPMS_7d_Expt/model_files/df_cleaned.rda")

df_cleaned %>%
  print(10)
```


Plot all data by mass and sample description on day 1:
```{r explore, fig.align = "center", fig.asp= 2, fig.width=8}
load("./model_files/df_cleaned.rda")

df_cleaned %>%
  filter(day == 1) %>%
  ggplot(aes(x = (cps + 0.1), y = factor(mass), color = description)) +
  geom_point() +
  scale_x_continuous(trans = "log10") +
  theme_cleveland()
```

Plot CPS @  m/z = 71.5 (Nd 2+), 75 (As), and 212 for each sample type.
```{r explore2, fig.align = "center", fig.asp= 1.25, fig.width=8}
df_cleaned %>%
  filter(mass %in% c(71.5, 75, 212)) %>%
  ggplot(aes(x = cps + 1, y = description, color = factor(mass), shape = factor(burn))) +
  geom_point(alpha = 0.5) +
  geom_jitter(height = 0.4) +
  scale_x_continuous(trans = "log10") +
  xlab("cps") +
  ylab("Sample ID") +
  theme_cleveland()
```

Plot CPS @ 9 m/z (Be) across days by burn replicate.
```{r explore3, fig.align = "center", fig.asp= 1, fig.width=8}
df_cleaned %>%
  filter(mass %in% c(9)) %>%
  ggplot(aes(x = description, y = cps + 1, color = factor(burn))) +
  geom_point(alpha = 0.5) +
  geom_jitter(width = 0.4) +
  facet_wrap(~ day, ncol = 2) +
  scale_y_continuous(trans = "log10") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  xlab("Sample ID") +
  ylab("log(cps + 1)") +
  theme_cleveland()
```

Plot CPS @ m/z = 212:231 in 0.8% $HNO_3$ blank solutions (samples A, B, C) by burn replicate.
```{r explore4, fig.align = "center", fig.asp= 0.5, fig.width=8}
df_cleaned %>%
  filter(description %in% c("0.8% HNO3 A", "0.8% HNO3 B", "0.8% HNO3 C"),
         mass %in% c(212:231)) %>%
  ggplot(aes(x = mass, y = cps, color = factor(burn))) +
  geom_point(alpha = 0.5) +
  #geom_jitter(width = 0.25) +
  facet_wrap(~ description) +
  #scale_y_continuous(trans = "log10") +
  xlab("mass") +
  ylab("cps")
```

```{r explore5, fig.align = "center", fig.asp= 0.5, fig.width=8}
df_cleaned %>%
  filter(description == "Multi Element",
         mass %in% c(212:231)) %>%
  ggplot(aes(x = mass, y = cps+1, color = factor(burn))) +
  geom_point(alpha = 0.5) +
  #geom_jitter(width = 0.25) +
  #facet_wrap(~ day) +
  #scale_y_continuous(trans = "log10") +
  xlab("mass") +
  ylab("cps")
```

Plot CPS @ m/z for 1/2 mass targets relevant to ideal background and according to $HNO_3$ sample and burn replicate. 
```{r explore6, fig.align = "center", fig.asp= 2, fig.width=8}
df_cleaned %>%
  filter(description %in% c("0.8% HNO3 A", "0.8% HNO3 B", "0.8% HNO3 C"),
         mass %in% c(6.5, 22.5, 23.5, 44.5, 58.5, 59.5, 62.5, 63.5, 65.5, 66.5, 69.5, 
                     70.5, 71.5, 72.5, 73.5, 74.5, 76.5, 77.5, 78.5, 81.5, 83.5, 91.5,
                     92.5, 99.5, 115.5, 120.5, 122.5, 139.5, 140.5, 141.5, 143.5, 144.5,
                     146.5, 147.5, 148.5, 151.5, 152.5, 159.5, 160.5, 204.5, 205.5, 212:229,
                     232.5)) %>%
  ggplot(aes(x = mass, y = cps, color = factor(burn))) +
  geom_point() +
  stat_smooth(method = "lm") +
  #geom_jitter(width = 1) +
  facet_wrap(~ day + description, ncol = 3) +
  #scale_y_continuous(trans = "log10") +
  xlab("mass") +
  ylab("cps")
```

Plot CPS @ m/z for 1/2 mass targets relevant to ideal background and according to $HNO_3$ sample for a single burn.
```{r explore7, fig.align = "center", fig.asp= 2, fig.width=8}
df_cleaned %>%
  filter(description %in% c("0.8% HNO3 A", "0.8% HNO3 B", "0.8% HNO3 C"),
         mass %in% c(6.5, 22.5, 23.5, 44.5, 58.5, 59.5, 62.5, 63.5, 65.5, 66.5, 69.5, 
                     70.5, 71.5, 72.5, 73.5, 74.5, 76.5, 77.5, 78.5, 81.5, 83.5, 91.5,
                     92.5, 99.5, 115.5, 120.5, 122.5, 139.5, 140.5, 141.5, 143.5, 144.5,
                     146.5, 147.5, 148.5, 151.5, 152.5, 159.5, 160.5, 204.5, 205.5, 212:229,
                     232.5),
         burn == 1) %>%
  ggplot(aes(x = mass, y = log(cps + 1), color = factor(burn))) +
  geom_point() +
  stat_smooth(method = "lm") +
  #geom_jitter(width = 1) +
  facet_wrap(~ day + description, ncol = 3) +
  #scale_y_continuous(trans = "log10") +
  xlab("mass") +
  ylab("cps")
```


Export csv file .

```{r summary_predict_X_export, eval=FALSE, include=TRUE}
df_cleaned %>%
  write_csv(file = "Individual_burn_data.csv")
```











